CLT Results

### 2022-04-11 analysis/clt_irt_analysis.Rmd
add bayesian model runs

```{r setup, include=FALSE}
library(tidyverse)
library(ggpubr)
library(tidyboot)
library(brms)
library(readxl)
library(ggthemes)
library(geomtextpath)

@@

@@

saveRDS(irt1, "../saves/irt1.rds")
```
### Model checks

@@

### TIC
```{r}
tic1 <- posterior_samples(irt1)%>% 
  select(b_Intercept, starts_with("r_targetWord"))%>%
  mutate(iter = 1:n()) %>% 
  pivot_longer(starts_with("r_targetWord"), names_to = "item", values_to = "xi") %>%
  mutate(item = str_extract(string = item, pattern = "(?<=\\[).*(?=,Intercept\\])"))%>%
  expand(nesting(iter, b_Intercept, item, xi),
         theta = seq(from = -6, to = 6, length.out = 100)) %>% 
  mutate(p = inv_logit_scaled(b_Intercept + xi + theta)) %>% 
  mutate(i = p * (1 - p)) %>% 
  group_by(theta, iter) %>% 
  summarise(sum_i = sum(i)) %>% 
  group_by(theta) %>% 
  summarise(i = median(sum_i))
```
```{r}
tic1%>%
  ggplot(aes(x = theta, y = i)) +
  geom_line() +
  labs(title = "TIC for the 1PL",
       x = expression(theta~('ability on the logit scale')),
       y = "Information") +
  theme_minimal()
```

### Differential Item Functioning

```{r}
irt_dif_dat <- data%>%
  filter(!is.na(sex))%>%
  select(subjID,sex, targetWord, correct)

irt1_f <- brm(
  data = irt_dif_dat%>%filter(sex == "w"),
  family = brmsfamily("bernoulli", "logit"),
  correct ~ 1 + (1 | targetWord) + (1 | subjID),
  prior = c(prior(normal(0, 1.5), class = Intercept),
            prior(student_t(10, 0, 1), class = sd)),
  cores = 4,
  chains = 4,
  iter = 8000
)

irt1_m <- brm(
  data = irt_dif_dat%>%filter(sex == "m"),
  family = brmsfamily("bernoulli", "logit"),
  correct ~ 1 + (1 | targetWord) + (1 | subjID),
  prior = c(prior(normal(0, 1.5), class = Intercept),
            prior(student_t(10, 0, 1), class = sd)),
  cores = 4,
  chains = 4,
  iter = 8000
)

```{r}
icc1_dif <- bind_rows(
  posterior_samples(irt1_m)%>%mutate(group = "m"),
  posterior_samples(irt1_f)%>%mutate(group = "f")
)%>%
  select(b_Intercept,group, starts_with("r_targetWord"))%>%
  group_by(group)%>%
  mutate(iter = 1:n()) %>% 
  ungroup()%>%
  pivot_longer(starts_with("r_targetWord"), names_to = "item", values_to = "xi") %>%
  mutate(item = str_extract(string = item, pattern = "(?<=\\[).*(?=,Intercept\\])"))%>%
  expand(nesting(iter,group, b_Intercept, item, xi),
         theta = seq(from = -6, to = 6, length.out = 100)) %>% 
  mutate(p = inv_logit_scaled(b_Intercept + xi + theta)) %>% 
  group_by(theta, group,item) %>% 
  summarise(p = mean(p))
```

```{r}
icc1_dif %>% 
  ggplot(aes(x = theta, y = p,group = group, col = group)) +
  geom_line(alpha = .75) +
  facet_wrap(~item)+
  #guides(col = F)+
  scale_color_viridis_d(name = "AoA") +
  labs(title = "Group Split by sex - 1PL",
       x = expression(theta~('ability on the logit scale')),
       y = expression(italic(p)(y==1))) +
  theme_minimal()
```

@@

```{r}
irt2 <- brm(
  data = irt_dat,
  family = brmsfamily("bernoulli", "logit"),
  bf(
    correct ~ exp(logalpha) * eta,
    eta ~ 1 + (1 |i| targetWord) + (1 | subjID),
    logalpha ~ 1 + (1 |i| targetWord),
    nl = TRUE
  ),
  prior = prior_2pl,
  control = list(adapt_delta = 0.9),
  cores = 6,
  chains = 6,
  iter = 6000
)

@@

saveRDS(irt2, "../saves/irt2.rds")
```

@@

@@

saveRDS(irt3_dif_sex, "../saves/irt3_dif_sex.rds")
```

@@

```{r}
coef_irt3_dif_sex <- coef(irt3_dif_sex)

eta_sex <- bind_rows(
  coef_irt3_dif_sex$targetWord[, , "eta_sexf"] %>%
	as_tibble(rownames = "item")%>%mutate(sex = "female"),
  coef_irt3_dif_sex$targetWord[, , "eta_sexm"] %>%
	as_tibble(rownames = "item")%>%mutate(sex = "male")
  )%>%mutate(nlpar = "Easiness")

alpha_sex <- 
  bind_rows(
    coef_irt3_dif_sex$targetWord[, , "logalpha_sexf"] %>%
	exp() %>%
	as_tibble(rownames = "item")%>%mutate(sex = "female"),
	    coef_irt3_dif_sex$targetWord[, , "logalpha_sexm"] %>%
	exp() %>%
	as_tibble(rownames = "item")%>%mutate(sex = "male")
  )%>%mutate(nlpar = "Discrimination")

params_sex <- bind_rows(eta_sex, alpha_sex) %>%
	select(-Est.Error) %>%
  left_join(aoa%>%rename(item = targetWord))

ggplot(params_sex,aes(reorder(item, -aoa_german_comb), Estimate, ymin = Q2.5, ymax = Q97.5)) +
	facet_wrap("nlpar", scales = "free_x") +
	geom_pointrange(aes(col = sex), position = position_dodge(width = .5)) +
	coord_flip() +
  scale_color_colorblind()+
	labs(x = "Item")+
  theme_minimal()
```

@@

saveRDS(irt3_dif_order, "../saves/irt3_dif_order.rds")
```

@@

```{r}
coef_irt3_dif_order <- coef(irt3_dif_order)

eta_order <- bind_rows(
  coef_irt3_dif_order$targetWord[, , "eta_orderorderA"] %>%
	as_tibble(rownames = "item")%>%mutate(order = "A"),
  coef_irt3_dif_order$targetWord[, , "eta_orderorderB"] %>%
	as_tibble(rownames = "item")%>%mutate(order = "B")
  )%>%mutate(nlpar = "Easiness")

alpha_order <- 
  bind_rows(
    coef_irt3_dif_order$targetWord[, , "logalpha_orderorderA"] %>%
      exp() %>%
      as_tibble(rownames = "item")%>%mutate(order = "A"),
    coef_irt3_dif_order$targetWord[, , "logalpha_orderorderB"] %>%
      exp() %>%
      as_tibble(rownames = "item")%>%mutate(order = "B")
  )%>%mutate(nlpar = "Discrimination")

params_order <- bind_rows(eta_order, alpha_order) %>%
	select(-Est.Error) %>%
  left_join(aoa%>%rename(item = targetWord))

ggplot(params_order,aes(reorder(item, -aoa_german_comb), Estimate, ymin = Q2.5, ymax = Q97.5)) +
	facet_wrap("nlpar", scales = "free_x") +
	geom_pointrange(aes(col = order), position = position_dodge(width = .5)) +
	coord_flip() +
  scale_color_colorblind()+
	labs(x = "Item")+
  theme_minimal()
```

@@

```{r}
remove <- c("telephone", "strawberry" ,"plane" ,"snake" ,"fish")

@@

firt_1F <- mirt(firt_dat, 1, itemtype = "2PL",  guess = .25,technical = list(NCYCLES = 1000))
firt_2F <- mirt(firt_dat, 2, itemtype = "2PL",  guess = .25,technical = list(NCYCLES = 1000))

anova(firt_1F, firt_2F)

@@

```{r}
traceline <- NULL
for(i in 1:43){
extr.2 <- extract.item(firt2, i)
Theta <- matrix(seq(-6,6, by = .1))
traceline[[i]] <- probtrace(extr.2, Theta)
}

@@

# plot chart
ggplot(format, aes(Theta, P.1, col = aoa_german_comb, group = item)) + 
  geom_line() + 
  #ggtitle('Probability Tracelines') + 
  xlab(expression(theta)) + 
  ylab(expression(P(theta))) + 
  #facet_wrap(~targetWord)+
  ylim(0,1)+
  scale_color_viridis_c(name = "AoA") +
  geom_hline(yintercept = 0.25, lty = 3, alpha = .75) +
  theme_bw() + 
  theme(
        axis.text.x=element_text(colour="black"),
        axis.text.y=element_text(colour="black"),
        legend.title=element_blank())
```

@@

sub <- data%>%
  filter(targetWord %in% select)%>%
  group_by(subjID)%>%
  summarise(mean_sub = mean(correct))

@@

ggplot(comp, aes(x = mean_full, y = mean_sub))+
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = .75)+
  geom_jitter(pch = 1, alpha = .75, height = 0.01, width = 0.01)+
  stat_cor()+
  ylim(0,1.02)+
  xlim(0,1.02)+
  theme_minimal()

@@

for (i in 1:1000) {

@@

>>>>

```{r setup, include=FALSE}
library(tidyverse)
library(ggpubr)
library(tidyboot)
library(brms)
library(readxl)
library(ggthemes)
library(geomtextpath)
library(coda)

@@

```{r}
estimate_mode <- function(s) {
  d <- density(s)
  return(d$x[which.max(d$y)])
}

hdi_upper<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","upper"])
}

hdi_lower<- function(s){
  m <- HPDinterval(mcmc(s))
  return(m["var1","lower"])
}
```

@@

saveRDS(irt1, "../saves/irt1.rds")

irt1 <- readRDS("../saves/irt1.rds")
```
### Model checks

@@

### TIC
```{r}
tic1 <- posterior_samples(irt1)%>% 
  select(b_Intercept, starts_with("r_targetWord"))%>%
  mutate(iter = 1:n()) %>% 
  pivot_longer(starts_with("r_targetWord"), names_to = "item", values_to = "xi") %>%
  mutate(item = str_extract(string = item, pattern = "(?<=\\[).*(?=,Intercept\\])"))%>%
  expand(nesting(iter, b_Intercept, item, xi),
         theta = seq(from = -6, to = 6, length.out = 100)) %>% 
  mutate(p = inv_logit_scaled(b_Intercept + xi + theta)) %>% 
  mutate(i = p * (1 - p)) %>% 
  group_by(theta, iter) %>% 
  summarise(sum_i = sum(i)) %>% 
  group_by(theta) %>% 
  summarise(i = median(sum_i))
```

```{r}
tic1%>%
  ggplot(aes(x = theta, y = i)) +
  geom_line() +
  labs(title = "TIC for the 1PL",
       x = expression(theta~('ability on the logit scale')),
       y = "Information") +
  theme_minimal()
```

@@

```{r}
irt2 <- brm(
  data = irt_dat,
  family = brmsfamily("bernoulli", "logit"),
  bf(
    correct ~ exp(logalpha) * eta,
    eta ~ 1 + (1 |i| targetWord) + (1 | subjID),
    logalpha ~ 1 + (1 |i| targetWord),
    nl = TRUE
  ),
  prior = prior_2pl,
  control = list(adapt_delta = 0.999, stepsize = 0.001, max_treedepth = 20),
  cores = 6,
  chains = 6,
  iter = 6000
)

@@

saveRDS(irt2, "../saves/irt2.rds")

irt2 <- readRDS("../saves/irt2.rds")
```

@@

irt3 <- readRDS("../saves/irt3.rds")
```

@@

saveRDS(irt3_dif_sex, "../saves/irt3_dif_sex.rds")

irt3_dif_sex <- readRDS("../saves/irt3_dif_sex.rds")
```

@@

```{r}
# eta_sex <- as_draws_df(irt3_dif_sex)%>%
#   select(b_eta_Intercept, starts_with("r_targetWord__eta"))%>%
#   mutate(iter = 1:n()) %>% 
#   pivot_longer(starts_with("r_targetWord")) %>%
#   mutate(name = str_remove(name, pattern = "r_targetWord__eta"),
#          name = str_remove_all(name, pattern = "\\[|\\]"))%>%
#   separate(name, into = c("item", "sex"), sep = "\\,")%>%
#   mutate(sex = str_remove(sex, pattern = "sex"),
#          val = b_eta_Intercept + value)%>%
#   group_by(item, sex)%>%
#   summarise(mode = estimate_mode(val),
#             uci = hdi_upper(val),
#             lci = hdi_lower(val))%>%
#   mutate(nlpar = "Easiness")
# 
# 
# alpha_sex <- as_draws_df(irt3_dif_sex)%>%
#   select(b_logalpha_Intercept, starts_with("r_targetWord__logalpha"))%>%
#   mutate(iter = 1:n()) %>% 
#   pivot_longer(starts_with("r_targetWord")) %>%
#   mutate(name = str_remove(name, pattern = "r_targetWord__logalpha"),
#          name = str_remove_all(name, pattern = "\\[|\\]"))%>%
#   separate(name, into = c("item", "sex"), sep = "\\,")%>%
#   mutate(sex = str_remove(sex, pattern = "sex"),
#          val = exp(b_logalpha_Intercept + value))%>%
#   group_by(item, sex)%>%
#   summarise(mode = estimate_mode(val),
#             uci = hdi_upper(val),
#             lci = hdi_lower(val))%>%
#   mutate(nlpar = "Discrimination")
# 
# params_sex <- bind_rows(eta_sex, alpha_sex) %>%
#   left_join(aoa%>%rename(item = targetWord))
# 
# saveRDS(params_sex, "../saves/model_params_irt3_dif_sex.rds")

params_sex <- readRDS("../saves/model_params_irt3_dif_sex.rds")

ggplot(params_sex,aes(reorder(item, -aoa_german_comb), mode, ymin = lci, ymax = uci)) +
	facet_wrap("nlpar", scales = "free_x") +
	geom_pointrange(aes(col = sex), position = position_dodge(width = .5)) +
	coord_flip() +
  scale_color_colorblind(labels = c("male","female"))+
	labs(x = "Item")+
  theme_minimal()
```

@@

saveRDS(irt3_dif_order, "../saves/irt3_dif_order.rds")

irt3_dif_order <- readRDS("../saves/irt3_dif_order.rds")
```

@@

eta_order <- as_draws_df(irt3_dif_order)%>%
  select(b_eta_Intercept, starts_with("r_targetWord__eta"))%>%
  mutate(iter = 1:n()) %>%
  pivot_longer(starts_with("r_targetWord")) %>%
  mutate(name = str_remove(name, pattern = "r_targetWord__eta"),
         name = str_remove_all(name, pattern = "\\[|\\]"))%>%
  separate(name, into = c("item", "order"), sep = "\\,")%>%
  mutate(order = str_remove(order, pattern = "orderorder"),
         val = b_eta_Intercept + value)%>%
  group_by(item, order)%>%
  summarise(mode = estimate_mode(val),
            uci = hdi_upper(val),
            lci = hdi_lower(val))%>%
  mutate(nlpar = "Easiness")

alpha_order <- as_draws_df(irt3_dif_order)%>%
  select(b_logalpha_Intercept, starts_with("r_targetWord__logalpha"))%>%
  mutate(iter = 1:n()) %>%
  pivot_longer(starts_with("r_targetWord")) %>%
  mutate(name = str_remove(name, pattern = "r_targetWord__logalpha"),
         name = str_remove_all(name, pattern = "\\[|\\]"))%>%
  separate(name, into = c("item", "order"), sep = "\\,")%>%
  mutate(order = str_remove(order, pattern = "orderorder"),
         val = exp(b_logalpha_Intercept + value))%>%
  group_by(item, order)%>%
  summarise(mode = estimate_mode(val),
            uci = hdi_upper(val),
            lci = hdi_lower(val))%>%
  mutate(nlpar = "Discrimination")

params_order <- bind_rows(eta_order, alpha_order) %>%
  left_join(aoa%>%rename(item = targetWord))

saveRDS(params_order, "../saves/model_params_irt3_dif_order.rds")

params_order <- readRDS("../saves/model_params_irt3_dif_order.rds")

ggplot(params_order,aes(reorder(item, -aoa_german_comb), mode, ymin = lci, ymax = uci)) +
	facet_wrap("nlpar", scales = "free_x") +
	geom_pointrange(aes(col = order), position = position_dodge(width = .5)) +
	coord_flip() +
  scale_color_colorblind()+
	labs(x = "Item")+
  theme_minimal()
```

@@

```{r}
remove <- c("telephone", "strawberry" ,"plane" ,"snake" ,"fish", "belt")

@@

firt_1F <- mirt(firt_dat, 1, itemtype = "2PL",  guess = .25,technical = list(NCYCLES = 2000))
firt_2F <- mirt(firt_dat, 2, itemtype = "2PL",  guess = .25,technical = list(NCYCLES = 2000))
firt_3F <- mirt(firt_dat, 3, itemtype = "2PL",  guess = .25,technical = list(NCYCLES = 2000))

anova(firt_1F, firt_2F)
anova(firt_1F, firt_3F)

@@

```{r}
traceline <- NULL
for(i in 1:46){
extr.2 <- extract.item(firt2, i)
Theta <- matrix(seq(-6,6, by = .1))
traceline[[i]] <- probtrace(extr.2, Theta)
}

@@

# plot chart
ggplot(format, aes(Theta, P.1, col = aoa_german_comb, group = item)) + 
  geom_line() + 
  #ggtitle('Probability Tracelines') + 
  xlab(expression(theta)) + 
  ylab(expression(P(theta))) + 
  facet_wrap(~targetWord)+
  ylim(0,1)+
  scale_color_viridis_c(name = "AoA") +
  geom_hline(yintercept = 0.25, lty = 3, alpha = .75) +
  theme_bw() + 
  theme(
        axis.text.x=element_text(colour="black"),
        axis.text.y=element_text(colour="black"),
        legend.title=element_blank())
```

@@

full <- data%>%
  group_by(age_group,subjID)%>%
  summarise(mean_full = mean(correct))

sub <- data%>%
  filter(targetWord %in% select)%>%
  group_by(age_group,subjID)%>%
  summarise(mean_sub = mean(correct))

@@

ggplot(comp, aes(x = mean_full, y = mean_sub))+
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = .75)+
  geom_jitter(pch = 1, alpha = .75, height = 0.01, width = 0.01)+
  stat_cor()+
  facet_wrap(~age_group)+
  ylim(0,1.02)+
  xlim(0,1.02)+
  theme_few()

@@

for (i in 1:200) {

@@

full <- data%>%
  group_by(age_group,subjID)%>%
  summarise(mean_full = mean(correct))

select2 <- dat%>%group_by(items)%>%
  summarise(mean_cor = mean(cor))%>%
  arrange(-mean_cor)%>%slice(1:15)%>%pull(items)

sub2 <- data%>%
  filter(targetWord %in% select2)%>%
  group_by(age_group,subjID)%>%
  summarise(mean_sub = mean(correct))

comp2<- full%>%left_join(sub2)

ggplot(comp2, aes(x = mean_full, y = mean_sub))+
  geom_abline(intercept = 0, slope = 1, lty = 2, alpha = .75)+
  geom_jitter(pch = 1, alpha = .75, height = 0.01, width = 0.01)+
  stat_cor()+
  facet_wrap(~age_group)+
  ylim(0,1.02)+
  xlim(0,1.02)+
  theme_few()
